---
title: "Four Parameters Categorical"
author: "Alec Rogers"
format: html
execute: 
  echo: false
---


```{r}
#| message: false

library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
```
Voting preferences may vary on the basis of sex, and understanding this relationship is crucial for designing effective campaign strategies. Using the nes data including information from the American National Election Studies survey, we investigate male and female voting preferences for the Bush, Clinton and Perot in the 1992 US presidential election. Our data is limited in that it comes from a survey: those who answer surveys may not be representative of voters in the 1992 US presidential election. We model voting as a multinomial function of voters' sex.

$$P(Y = k) = \frac{e^{\beta_{k0} + \beta_{k1} X_1 + \beta_{k2} X_2 + \cdots + \beta_{kn} X_n}}{\sum_{j=1}^{K} e^{\beta_{j0} + \beta_{j1} X_1 + \beta_{j2} X_2 + \cdots + \beta_{jn} X_n}}$$

with $Y \sim \text{Multinomial}(\boldsymbol{\rho})$ where $\boldsymbol{\rho} = (\rho_1, \rho_2, \ldots, \rho_K)$ are the probabilities above.


Now we set up our data set
```{r}
#| cache: true


nes_92 <- nes |> 
  filter(year == 1992) |> 
  select(sex, pres_vote) |> 
  drop_na() |> 
  mutate(pres_vote = as.factor(case_when(
    pres_vote == "Democrat" ~ "Clinton",
    pres_vote == "Republican" ~ "Bush",
    pres_vote == "Third Party" ~ "Perot",
  ))) 

fit_nes <- multinom_reg(engine = "nnet") |>
  fit(pres_vote ~ sex, data = nes_92)
```

Possible $\LaTeX$ representations of this model
$$
Y_i \in \{\text{Clinton}, \text{Bush}, \text{Perot}\}
$$

$$
Y_i \mid \text{sex}_i \sim \operatorname{Categorical}(\boldsymbol{\pi}_i)
$$

$$
\boldsymbol{\pi}_i =
\left(
\pi_{i,\text{Clinton}},\,
\pi_{i,\text{Bush}},\,
\pi_{i,\text{Perot}}
\right),
\qquad
\sum_{k} \pi_{i,k} = 1
$$

Let Bush be the baseline category. Then the model is:

$$
\log\left(\frac{\pi_{i,\text{Clinton}}}{\pi_{i,\text{Bush}}}\right)
= \beta_{0,\text{Clinton}} + \beta_{1,\text{Clinton}} \cdot \text{sex}_i
$$

$$
\log\left(\frac{\pi_{i,\text{Perot}}}{\pi_{i,\text{Bush}}}\right)
= \beta_{0,\text{Perot}} + \beta_{1,\text{Perot}} \cdot \text{sex}_i
$$

Another representation

$$
\pi_{i,\text{Bush}} = 1 - \pi_{i,\text{Clinton}} - \pi_{i,\text{Perot}}
$$

```{r}
library(dplyr)
library(gt)

fit_nes |>
  tidy(conf.int = TRUE) |>
  mutate(
    term = recode(term,
      `(Intercept)` = "Intercept",
      sexMale = "Male (vs Female)"
    ),
    y.level = recode(y.level,
      Clinton = "Clinton vs Bush",
      Perot   = "Perot vs Bush"
    )
  ) |>
  select(
    Outcome = y.level,
    Term = term,
    Estimate = estimate,
    `Std. Error` = std.error,
    `CI Lower` = conf.low,
    `CI Upper` = conf.high
  ) |>
  gt() |>
  fmt_number(
    columns = c(Estimate, `Std. Error`, `CI Lower`, `CI Upper`),
    decimals = 2
  ) |>
  cols_label(
    Outcome = "Outcome (Baseline: Bush)"
  ) |>
  tab_header(
    title = "Multinomial Logistic Regression Results",
    subtitle = "1992 U.S. Presidential Vote by Sex"
  )
```
